---
title: "Olink Reference Normalization"
output: html_document

params:
  rp_upload_data: NA
  rp_bridging_str: NA
  rp_norm_method: NA
  rp_combined_data: NA
  rp_combined_meta: NA
---

```{r eval = FALSE, echo = FALSE}
# For PDF output, change the header to have "output: pdf_document".
#
# Note that due to an issue in rmarkdown, the default value of a parameter in
# the header cannot be `NULL`, so I used a default of `NA` for the default value
# of `n`.
library(dplyr)
library(DT)
```


## Summary of plate level information.


```{r, echo=F, message=F, warning=F}
DT::datatable(
      data.frame(software_version = sapply(params$rp_upload_data, function(x) x@metadata$software_version) %>% unlist(),
                 filename = sapply(params$rp_upload_data, function(x) x$file_name[1]) %>% unlist(),
                 panel = sapply(params$rp_upload_data, function(x) x@metadata$panel)%>% unlist(),
                 total_samples = sapply(params$rp_upload_data, function(x) ncol(x)) %>% unlist(),
                 total_analyte = sapply(params$rp_upload_data, function(x) nrow(x)) %>% unlist()),
      rownames = NULL,
      options = list(scrollX = TRUE))
```
 
### Reference sample identifier: `r paste(strsplit(params$rp_bridging_str, split = ","), collapse = ",")`    

```{r, echo=F, message=F, warning=F}
# needs sample_id, Plate.ID, in metafile and "lot" only appear once in column names

lapply(strsplit(params$rp_bridging_str, split = ",")%>%unlist(), function(x){
      bdg_ls <- pull_bdg(params$rp_upload_data, pattern = x)
      data.frame(filename = names(bdg_ls),
                 identifier = x,
                 total_samples = sapply(bdg_ls, function(x) ncol(x)) %>% unlist(),
                 total_analyte = sapply(bdg_ls, function(x) paste(x$Assay, collapse = ",")) %>% unlist())
    })%>%
      do.call(what = "rbind")%>%
      DT::datatable(rownames = NULL,
      options = list(scrollX = TRUE))

```

### Normalization Method:
The plate-to-plate variation was adjusted based on adjust-factor(plate specific) which makes equal NPX value among reference samples identified above.    

Between Plate Normalization Method: **`r paste(params$rp_norm_method)`**    


### PCA Before Normalization 

```{r, echo=F, out.width= "100%", message=F, warning=F}
mat <- params$rp_combined_data[ ,!grepl("(^NC|^IPC|randox)", ignore.case = T, params$rp_combined_data$Assay)]
ref_id <- strsplit(params$rp_bridging_str, split = ",")%>%unlist()

fit <- prcomp(impute::impute.knn(t(mat@assays@data$npx))$data)
idf <- paste(ref_id, collapse = "|")
idf <- paste0("(", idf, ")")

temp <- data.frame(pc_1 = fit$x[ ,1], #fit$layout[ ,1],
           pc_2 = fit$x[ ,2],#fit$layout[ ,2],
           sample_id = mat$Assay,
           type = ifelse(grepl(idf, ignore.case = T, mat$Assay), "Ref", "Study_Sample"),
           Plate.ID = trim_string_bycommon(mat$Plate.ID, split = "_"))
  
temp %>%
  ggplot()+
  geom_point(aes(pc_1, pc_2, color = Plate.ID, shape = type))+
  geom_line(data = temp%>%
              filter(sample_id %in% ref_id),
            aes(pc_1, pc_2, group = sample_id), linetype = 2, alpha = 0.6)+
  facet_wrap(~type)+
  theme_bw()+
  theme(legend.position="bottom", legend.box="vertical", legend.margin=margin())
```

### PCA After Normalization 

```{r, echo=F, out.width= "100%", message=F, warning=F}
#mat <- impute::impute.knn(t(params$rp_combined_data@assays@data$normed))$data

fit <- prcomp(impute::impute.knn(t(mat@assays@data$normed))$data)
idf <- paste(ref_id, collapse = "|")
idf <- paste0("(", idf, ")")

temp <- data.frame(pc_1 = fit$x[ ,1], #fit$layout[ ,1],
           pc_2 = fit$x[ ,2],#fit$layout[ ,2],
           sample_id = mat$Assay,
           type = ifelse(grepl(idf, ignore.case = T, mat$Assay), "Ref", "Study_Sample"),
           Plate.ID = trim_string_bycommon(mat$Plate.ID, split = "_"))

temp %>%
  ggplot()+
  geom_point(aes(pc_1, pc_2, color = Plate.ID, shape = type))+
  geom_line(data = temp%>%
              filter(sample_id %in% ref_id),
            aes(pc_1, pc_2, group = sample_id), linetype = 2, alpha = 0.6)+
  facet_wrap(~type)+
  theme_bw()+
  theme(legend.position="bottom", legend.box="vertical", legend.margin=margin())
```

### Reference Sample QC Plot

**Black dots indicate LOD**       

```{r, echo=F, out.width= "100%", message=F, warning=F, fig.height = 15}
#ref_sample <- strsplit(params$rp_bridging_str, split = ",")%>%unlist()
plot_npx_norm_qc(params$rp_combined_data, bridge_pattern = idf)$all
```

